Xor-Some Preregistered Study: Preprocessing

### 2022-03-31 analysis/main/01_xor-some_prereg_preprocessing_final.Rmd
added comments on preprocessing

```{r, echo=FALSE}
d_critical_clean <- d_full_clean %>% filter(condition == "critical")
d_critical_long <- d_critical_clean %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")

d_critical_long <- d_critical_long %>% 
  mutate(block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
```

Create more extensive condition labels.

``` {r}
# extending 'conditions' labels to include whether the utterance was present or not
d_critical_long <- d_critical_long %>% 
  mutate(
    block_extended =  ifelse(block %in% c("some", "xor"), "target", block)
  )
```

Z-scoring:
```{r}
d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))

d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )

>>>>

```{r, echo=FALSE}
d_critical_clean <- d_full_clean %>% filter(condition == "critical") %>% 
  mutate(
    block = case_when(
      block == "comp" ~ "competence", 
      block == "rel"  ~ "relevance", 
      block == "pri"  ~ "prior",
      T ~ block) ,
    block_extended = case_when(
      block == "some" ~ "target", 
      block == "xor"  ~ "target", 
      T ~ block) )

# write_csv(d_critical_clean, "../../data/main/results_prereg_final_tidy-MF.csv")

# MF: Careful! This step introduces tripling each individual `response` measure

d_critical_long <- d_critical_clean %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")

Z-scoring:

MF: Here is a problem. The `d_critical_long` contains every `response` measure three times.
    This is unproblematic for the mean, but not of the SD.

```{r}
# MF: ooriginal code

d_critical_zScored <- d_critical_long %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))
```

MF: To see this problem, compare the following (the first is wrong and what we used so far, the second correct):

```{r}

d_critical_long %>% 
  select(submission_id, block_extended, response) %>% 
  group_by(submission_id, block_extended) %>%
  summarize(mean = mean(response),
            sd = sd(response)) %>% 
  mutate(data_used = 'long') %>% 
  arrange(submission_id, block_extended)
```

```{r}
d_critical_clean %>% 
  select(submission_id, block_extended, response) %>% 
  group_by(submission_id, block_extended) %>%
  summarize(mean = mean(response),
            sd = sd(response)) %>% 
  mutate(data_used = 'short') %>% 
  arrange(submission_id, block_extended)
```

We should therefore use the following (but now the table doesn't contain the high/low classifications as individual rows):

```{r, eval = F}
# MF: corrected code; not executed

d_critical_zScored <- d_critical_clean %>% group_by(submission_id, block_extended) %>%
  mutate(block_mean = mean(response),
         block_sd = sd(response),
         response_centered = (response - block_mean)/block_sd,
         # catch the cases where sd is 0 
         response_centered = ifelse(is.na(response_centered), 0, response_centered))
```

Wrangling into wide format.

```{r}
d_critical_zScored_wide <- d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered) %>% 
  unique() %>% 
  pivot_wider(
    names_from = block_extended, 
    values_from = response_centered, 
    values_fn = mean # getting means for double prior measurement in "xor"
  )
